---
title: "Shiny training intro- session 3"
author: "Chris Beeley"
date: "February 2021"
output: 
  ioslides_presentation:
    css: temp.css
runtime: shiny
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## Recap

* Reactive expressions
* Tab panels

## Dynamic UI

* Sometimes you need your UI to react to the state of your application
* You can run `unique(ShinyContactData$Year)` in the UI definition
* But you can't access reactive values
    * e.g. `unique(returnData()$Year)`

## Example

<iframe src="https://involve.nottshc.nhs.uk:8443/experience"></iframe>

## Code

```{r, eval = FALSE}

# ui

uiOutput("trustControl")

# server

output$trustControl <- renderUI({
  
  selectInput("trust",
              "Select Trust",
              choices = unique(ae_attendances$Name),
              multiple = TRUE)
})

```

## Exercise

* We're going to access the reactive object within the UI and only display the relevant data
* This example is a little contrived because we need more code for a real example (easily 100 lines in the example I showed)
* Open sitrep_third in session three

## Validating state of application

* Shiny has two ways of avoiding these
* req()
* req() will refrain from running a block of code until whatever is inside it takes a valid value
* validate() will also refrain from running a block of code but will also write a nice error message

```{r eval = FALSE}

# req()

req(input$trust) # this will not run until input$trust has a value

# validate()

validate(
  need(input$trust, "Please select a trust"),
  need(input$somethingElse "Please select a value for this input too")
)

```

## Add req/ validate to the application

* Keep on the same application and make a copy if you want to keep the one you just edited
* Run the code and see what happens when you deselect filters when the graph is visible
* Again keep the previous application so you have a copy
* Let's fix the error message together

## The layout functions

* There are LOTS of layout functions other than just sidebarLayout
* flowLayout(), fluidRow(), navBar(), navList(), splitLayout(), verticalLayout()
* We'll cover the fundamental one and then do others if we have time

## fluidRow()

```{r eval = FALSE}

ui <- fluidPage(
  fluidRow(
    column(width = 6, 
           wellPanel(p("This appears in one column"))
    ),
    column(width = 6, 
           wellPanel(p("Then this in another column"))
    )
  ),
  fluidRow(
    column(width = 3, 
           wellPanel(p("Then a thin column underneath the top row"))),
    column(width = 9, 
           wellPanel(p("Then a wide column next to that one")))
  )
)
server <- function(input, output) {}
shinyApp(ui = ui, server = server)

```

## Example

```{r}

shinyAppDir("layout_example/")

```

## Exercise

* Keep the same application (copy it if you want to)
* We're going to redo it with the fluidRow() and column() functions

## Bonus material

* Try some of the others
* There are [loads of examples online](https://github.com/ChrisBeeley/Web-Application-Development-with-R-Using-Shiny-third-edition/tree/master/Chapter04)
* You can have a go now if there's time or in your own time another time
* Please fill out the evaluation (your employers want to see it 😄)
