---
title: "Shiny training- session 2"
author: "Chris Beeley"
date: "November 2020"
output: 
  ioslides_presentation:
    css: temp.css
runtime: shiny
---

```{r setup, include=FALSE}

library(DiagrammeR)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## Recap

* Application setup
* Reactivity

## Reactive expressions

- Reactive expressions are expressions that change when their inputs change
- They enable you to maintain the data fetch/ clean code for several outputs in one place
- Crucially, they *cache* their results, which means that it's run once, for all the outputs, and only changes if its inputs change
- Great for computationally intensive or bandwidth heavy data functions

## Standard Shiny

```{r, echo = FALSE, warning = FALSE}

node_df <- create_node_df(n = 5 ,
                          label = c("Input 1", "Input 2", "Output 1", 
                                    "Output 2", "Output 3"),
                          fontsize = 24,
                          fillcolor = c("green", "green", "red", "red", "red"),
                          fontcolor = "black")

edge_df <- data.frame("from" = c(1, 1, 2, 2),
                      "to" = c(3, 4, 4, 5))

i_graph_1 <-
  create_graph()

i_graph_2 <-
  i_graph_1 %>%
  add_nodes_from_table(
    table = node_df,
    label_col = label)

i_graph_3 <-
  i_graph_2 %>%
  add_edges_from_table(
    table = edge_df,
    from_col = from,
    to_col = to,
    from_to_map = id_external)

i_graph_3 %>% render_graph()

```

## With reactive object

```{r, echo = FALSE, warning = FALSE}

node_dfR <- create_node_df(n = 6,
                           label = c("Input 1", "Input 2", "Reactive object", "Output 1", 
                                     "Output 2", "Output 3"),
                           fontsize = 24,
                           fillcolor = c("green", "green", "blue", "red", "red", "red"),
                           fontcolor = "black")

edge_dfR <- data.frame("from" = c(1, 2, 3, 3, 3),
                       "to" = c(3, 3, 4, 5, 6))

i_graph_1 <-
  create_graph()

i_graph_2 <-
  i_graph_1 %>%
  add_nodes_from_table(
    table = node_dfR,
    label_col = label)

i_graph_3 <-
  i_graph_2 %>%
  add_edges_from_table(
    table = edge_dfR,
    from_col = from,
    to_col = to,
    from_to_map = id_external)

i_graph_3 %>% render_graph()

```

## Add reactive expression

* sitrep_second.R contains all server code
* sitrep_second folder contains the answer

## Debugging

* Start simple
* Make sure your code works outside of Shiny!
* Use of cat()
* Breakpoints and browser()
* Run options(shiny.reactlog = TRUE)
    * Run
    * Press Ctrl+F3 for an interactive summary of reactive activity in your application
    * This has saved me more than once

## Tabset panels

* Tabset panels are useful if you want multiple outputs but not all on the same page
* If you give them an id you can test for which is selected too which is very useful.

```{r, eval = FALSE}

mainPanel(
  tabsetPanel(id = "tabset",
              tabPanel("Label", outputFunction("nameOfOutput")),
              tabPanel("Label2", outputFunction2("nameOfOutput2"))
  )
)

```

## Add tab panel

* We'll add a tab panel to the application
* Keep the previous one so you have both versions
* What happens when you deselect a status or year?
* Now we're going to fix this

## Validating state of application

* Shiny has two ways of avoiding these
* req()
* req() will refrain from running a block of code until whatever is inside it takes a valid value
* validate() will also refrain from running a block of code but will also write a nice error message

```{r eval = FALSE}

# req()

req(input$trust) # this will not run until input$trust has a value

# validate()

validate(
  need(input$trust, "Please select a trust"),
  need(input$somethingElse "Please select a value for this input too")
)

```

## Add req/ validate to the application

* Again keep the previous application so you have a copy
